\chapter{Dired}

Entering Dired mode with \verb|C-x C-f| adding a directory.

\section{Executing Commands in Shell Buffers}

\begin{description}
\item[M-!] Execute command string in inferior shell; display output, if any. With \verb|C-u| prefix, insert the output into the current buffer.
\item[M-x shell] Run an inferior shell.
\item[M-|] Run shell command on region.
\end{description}

You can have multiple shell buffers running at once; just use the command \verb|M-x rename-uniquely| to rename your shell buffer.
Then type \verb|M-x shell| to open a new one.



\section{Operations}
\begin{description}
\item[v] View file read only, q to quit.
\item[RET] Find the file, \verb|C-x b| to get back.
\item[x] Delete the deleted-flagged files.
\item[D] Do delete.
\item[C] Do copy.
\item[R] Do rename.
\item[Z] Compress or uncompress marked or next ARG files.
\item[=] Run the \verb|diff| command on the given two files.
\item[!] Run shell commond.
\item[g] Replace current buffer text with the text of the visited file on disk.
\item[+] Create a directory.
\item[Q] Replace matches of FROM with TO, in all marked files.
\item[A] Find all matches for REGEXP in all marked files.
\end{description}

\section{Marking}

\begin{description}
\item[m] Mark the current file.
\item[3m] Mark the current file and the next 2 files.
\item[M-Del] Unmark all files.
\item[U] Unmark all files.
\item[t] Toggle marks.
\item[* *] Mark excutable files.
\item[* @] Mark symlinks.
\item[* /] Mark directories.
\item[\% m] Mark files matching REGEXP.
\item[\% g] Mark files containing REGEXP.
\item[d] Flag file as deletion.
\item[\#] Flag autosave files.
\item[\~{}] Flag backup files.
\item[u] Unmark the current file.
\item[Del] Unmark backward.
\end{description}


\section{Navigating}

\begin{description}
\item[\^{}] Up directory.
\item[>] Next directory line.
\item[<] Previous directory lien.
\item[i] Insert this subdirectory into the same dired buffer.
\end{description}

