
\chapter{质量}

\section{错题}

1、 [单选] 某新聘项目经理接手了一个正在开展的项目。审查项目管理计划时，该项目经理意识到，质量管理计划偏离了已知的质量要求。项目经理首先应该做什么？

A newly hired project manager took over an ongoing project. When reviewing the project management plan, the project manager realized that the quality management plan deviated from known quality requirements. What should the project manager do first?

A：与项目相关方一起审查质量管理计划

Review the quality management plan together with project stakeholders

B：准备并提交更新质量管理计划的变更请求

Prepare and submit a change request to update the quality management plan

C：确保项目遵循当前的质量管理计划

Ensure that the project follows the current quality management plan

D：将该项目制定为质量标准可发生变化的敏捷项目

Develop projects as agile projects with good quality criteria

正确答案：B

你的答案：A

解析：解析：PMBOK（6）P306-8.3.3.4控制质量-变更请求。这道题目容易选A，但是要注意“审查项目管理计划时”，说明选项A的内容已经做了，那么接下来针对发现的问题进行纠正，所以选B。选项C，管理计划已经不可用。选项D，情景中没有提到质量标准灵活多变，所以采用敏捷的方式进行管理欠妥。


2、 [单选] 质量保证团队发现某制造项目未能交付优质产品，但项目团队成员拒绝接受这些调查结果。项目经理应该如何减轻这一冲突？

The quality assurance team found that a manufacturing project failed to deliver high-quality products, but the project team members refused to accept these investigations. How should the project manager mitigate this conflict?

A：与项目团队一起审查质量管理计划

Review the quality management plan with the project team together

B：推动项目团队与质量保证团队开展富有成效的协作

Drive a productive one between the project team and the quality assurance team

C：继续实施质量管理计划

Continue to implement the quality management plan

D：要求项目团队接受质量保证调查结果

Request the project team to accept quality assurance

正确答案：A

你的答案：B

解析：解析：PMBOK（6）P288，8.2-管理质量。质量保证团队和质量控制团队对质量活动没有达成统一意见，此时需要对质量管理计划进行审查。因为问题出在对质量管理活动的分歧上，所以选项B将其视为冲突进行管理欠佳，排除。选项C，在质量管理计划获得认可前继续实施不现实。选项D，过于武断。



5、 [单选] 一所大学有一个项目，为新录取的学生实施新的在线申请系统，试验解决方案显示，大量申请捕捉到不完整和不准确的数据。项目团队、主题专家(SME)和关键相关方召开了一次研讨会，以调查这些错误的原因。若要专注于正确的领域进行变更，从而提高数据捕捉流程的质量，项目经理应该与团队一起使用下列哪一项技术？

A university has a project to implement a new online application system for new students, and pilot solutions show that large numbers of applications capture incomplete and inaccurate data. The project team, subject matter experts (SME), and key stakeholders held a workshop to investigate the causes of these errors. To focus on making changes in the right areas to improve the quality of the data capture process, which of the following techniques should the project manager use with the team?

A：因果图

Cause and effect diagrams

B：亲和图

Affinity diagrams

C：影响图

Influence diagrams

D：鱼骨图

Fishbone diagrams

正确答案：B

你的答案：D

解析：解析：PMBOK（6）P293，8.2.2.4数据表现-亲和图。亲和图可以对潜在缺陷成因进行分类，展示最应关注的领域。这道题目容易选A或者D，但是要注意“大量申请捕捉到不完整和不准确的数据”，首先需要对捕捉到的数据进行分类才能展开分析。选项C，影响图是不确定条件下决策制定的图形辅助工具。它将一个项目或项目中的一种情境表现为一系列实体、结果和影响，以及它们之间的关系和相互影响。


15、 [单选] 质量需求已被识别，项目正在测试和评估这些需求。项目经理现在需要将测试结果与质量需求进行对比，以帮助确保这些结果会被接受。 项目经理接下来该做什么？

Quality requirements have been identified and the project is testing and evaluating these requirements. Project managers now need to compare test results with quality requirements to help ensure that these results are accepted. What should the project manager do next?

A：风险审查

The risk review

B：质量报告

The quality report

C：数据分析报告

The data analysis report

D：质量审查

The Quality review

正确答案：D

你的答案：B

解析：解析：PMBOK（6）P298，8.3-控制质量。“将测试结果与质量需求进行对比”，属于控制质量活动。选项A，风险审查，属于风险管理工具。选项B是控制质量过程的的输出。选项C是数据分析后的输出。


19、 [单选] 某银行财务部门负责人联系项目经理说，该行现有软件需要更新，以便将一组新需求包含进来。因此，项目经理正在实施该应用程序的一个新版本，项目经理查阅了一份可用资源名单，并确定应聘请专家，以便核实所有质量标准是否均已达到。项目经理应首先向专家提供哪个文件以便核实质量标准？

The head of a bank's finance department contacted a project manage, claiming the bank's existing software needed to be updated to include a new set of requirements. As a result, the project manager is implementing a new version of the application and has reviewed a list of available resources and determined that an expert should be hired to verify that all quality criteria have been met. Which document should the project manager first provide to the expert to verify the quality criteria?

A：质量审计

The Quality audit

B：质量管理计划

The Quality management plan

C：问题日志

The Issue log

D：风险登记册

The risk register

正确答案：B

你的答案：A

解析：解析：PMBOK（6）P286，8.1.3.1-质量管理计划。质量标准记录在质量管理计划当中。选项A，质量审计属于工具。选项C，问题日志记录已发现的问题，侧重于已发现问题的聚焦和跟踪。选项D，风险登记册用于记录已识别的不确定性事件。


20、 [单选] 在项目执行期间，监管合规团队公布了新的质量标准。更新质量管理计划时，项目经理应考虑什么？

During the execution of the project, the regulatory compliance team published new quality criteria. What should the project manager consider when updating the quality management plan?

A：经验教训登记册

The Lessons learned register

B：成本效益分析

Cost-benefit analysis

C：风险登记册

The risk register

D：事业环境因素

Enterprise environmental factors

正确答案：D

你的答案：C

解析：解析：PMBOK（6）P301，8.3.1.6-控制质量-事业环境因素。“新的质量标准”是对当前项目的制约，所以属于事业环境因素。选项A，经验教训登记册记录在项目实施期间识别的经验教训。选项B，成本效益分析是对财务投入和产出关系的分析。选项C，风险登记册记录已识别的风险。



21、 [单选] 在某项目中，需要为某连锁药店生产一系列健康补充品，客户向项目经理报告称某些产品未达到质量要求，他们似乎存在同样的缺陷。项目经理该做什么？

In a project to produce a range of health supplements for a chain of pharmacies, the customer reported to the project manager that some products did not meet quality requirements and that they appeared to have the same defects. What should the project manager do?

A：加强检查

Strengthen inspection

B：进行根本原因分析

Execute root cause analysis

C：审视各个过程

Review all processes

D：增加统计抽样

Increase statistical sampling

正确答案：C

你的答案：B

解析：解析：PMBOK（6）P294，8.2.2.5-审计。这道题目很容易选B。注意情景描述，“似乎存在同样的缺陷”，属于大面积出现的共性问题，此时要从流程上和标准上去分析导致质量问题发生的根源。如果针对的是孤立的个案，宜采用根本原因分析。选项AD是在解决当前后可能做出的调整，这些质量控制活动要更新到质量管理计划。


22、 [单选] 虽然已为一个多阶段项目制定了项目管理程序,但在第一阶段产生了质量较差的可交付成果。项目经理首先应该做什么？

While project management procedures have been developed for a multi-phase project, the first phase has resulted in poor quality deliverables. What should the project manager do first?

A：识别根本原因

Identify the root cause

B：审计质量过程

Audit quality processes

C：更新质量管理过程

Update the quality management process

D：增加完成质量控制的资源

Increase resources to complete quality control

正确答案：A

你的答案：B

解析：解析：PMBOK（6）P303，8.3.2.2-数据分析-根本原因分析。因为情景中没有强调是大面积的共性问题，所以此时不采用质量审计，排除B选项。选项CD，是在解决当前问题后可能做出的调整。


27、 [单选] 由于产品的重要性，发起人希望密切监督项目的进展情况，以确保质量超出市场标准。项目经理应该做什么来满足发起人的需求？

Due to a product’s high importance, the sponsor wants to closely monitor the project's schedule to ensure that the quality exceeds market criteria. What should the project manager do to meet the sponsor’s requirements?

A：应用关键路径法

Apply the critical path method

B：使用项目报告

Use project reporting

C：评估时间提前量和时间滞后量

Evaluate leads and time lags

D：执行数据分析

Execute data analysis

正确答案：D

你的答案：B

解析：解析：PMBOK（6）P303-8.3.2.2，数据分析。题干中的关键字：“确保质量超过市场标准”。根据题目关键字需要只有用控制质量来保证产品的质量能够达到标准，那么选项中选择和控制质量相关的内容，选项D正确。关键路径，提前量和滞后量都是和进度相关和质量无关，排除AC。项目报告包括太多内容，不够精确，排除B。

