
\chapter{进度}

\section{错题}

2、 [单选] 项目经理正在估算其组织之前未从事过的一类项目的成本.应该使用什么方法来估算成本？

The project manager is estimating the cost of a type of project that his organization has not undertaken before. What method should be used to estimate the cost?

A：参数估算

Parametric estimating

B：自下而上估算

Bottom-up estimating

C：自上而下估算

Top-down estimating

D：三点估算

Three-point estimating

正确答案：B

你的答案：D

解析：解析：PMBOK（6）P202，6.4.2.5-自下而上估算。“从未从事过的一类项目”，没有可以用以参照的项目，排除参数估算和类比估算。选项D，三点估算在估算的过程中兼顾风险因素，情景中没有强调类似的场景，排除。



4、 [单选] 在进行绩效评审后，项目经理识别到关键路径上五个工作包的其中一个可能会延迟两周时间。若要确保项目按进度执行，项目经理应该做什么？

In conducting the report, the project manager identified one of the five work packages on the critical path that could be delayed by two weeks. What should the project manager do to ensure that the project is on schedule?

A：通知相关方，并请求更多时间来处理该工作包

Notify the stakeholders and request more time to process the work package

B：请求额外的资源来处理受影响的工作包

Request additional resources to handle the affected work packages

C：分析该工作包以确定是否可以使用一些浮动时间

Analyze the work package to determine if some floating time is available

D：调整受影响工作包的进度管理计划

Adjust the schedule management plan for the affected work packages

正确答案：B

你的答案：C

解析：解析：PMDOK（6）P215，6.5.2.6-进度压缩。选项A，焦点是工期出现了延迟，再请求更多的时间，问题更加严重。选项C，“关键路径”，浮动时间为0。选项D，进度管理计划是针对整个项目的，而不是针对工作包，错误。


6、 [单选] 由于监管要求，项目的上线日期是固定的，项目经理在规划阶段运行了关键路径这种方法，结果显示该项目预计会比规定时间晚两个月完成。项目经理该做什么来纠正这种情况？

Due to regulatory requirements, the project's on-line date is fixed, and the project manager runs a critical path approach during the planning phase, which shows that the project is expected to be completed two months later than the required time. What should the project manager do to correct the project?

A：更新成本管理计划，为该项目分配更多资源

Update the cost management plan to allocate more resources to the current project

B：包含进度计划储备或紧急情况，并分析新的关键路径

Contain schedule reserves or contingencies, and analyzes new critical paths

C：进行反向排程，以重新定义所需的时间和资源

Reverse scheduling to redefine the time and resources required

D：通过剔除某些活动和缩短期限来缩减项目范围

Eliminate the elimination of certain activities and reduce deadlines to reduce the project scope

正确答案：C

你的答案：B

解析：解析：PMBOK（6）P209-6.5.2.1进度网络分析。“上线的日期是固定的”，所以要以上线日期进行分析，根据分析的结果调整项目管理计划，譬如缩小项目范围、追加资源投入等等。选项ABD在选项C之后。



8、 [单选] 项目经理必须确定项目的完成日期。虽然大多数活动都是已知的，但由于产品交付日期未知，某些工作包无法安排。项目经理应该使用什么来向相关方提供完工日期估算？

The project manager must determine the project finish date. Although most activities are known, some work packages cannot be scheduled because the product delivery date is unknown. What should the project manager use to provide the stakeholder with an estimate of the finish date?

A：滚动式规划

Rolling wave planning

B：蒙特卡洛模拟

Monte Carlo simulation

C：分解

Decomposition

D：计划评审技术（PERT)

Program Evaluation and Review Technique (PERT)

正确答案：B

你的答案：D

解析：解析：PMBOK（6）P213-6.5.2.4数据分析-模拟。这题易选错成A。“必须确定项目的完工日期”，采用滚动式规划就不合适了，排除AC选项。此时需要借助蒙特卡洛技术，对项目完成日期的可能分布进行预测。



16、 [单选] 负责跟踪项目文件的团队成员被指派到另一个项目，项目经理现在正在寻求项目文件，以便帮助制定进度计划。项目经理应使用哪些项目文件？

The team member responsible for tracking project documentation has been assigned to another project, and the project manager is now seeking project documentation. To help make schedule plans, what project documents should the project manager use?

A：经验教训登记册、里程碑列表、项目团队任务分配表

Lessons learned register, milestones list, project team task assignment list

B：里程碑列表、资源需求和项目章程

List of milestones, resource requirements, and project charter

C：范围基准、里程碑列表和活动清单

Scope baselines, milestone lists, and activity lists

D：活动清单、假设日志和进度基准

Activity lists, assumption logs, and schedule baselines

正确答案：C

你的答案：B

解析：解析：PMBOK（6）P207-6.5.1制定进度计划：输入。 题干关键词“帮助制定进度计划”。 选项A错误，项目团队任务分配表，即责任分配矩阵，具体到每个成员对每个工作包的职责。而项目团队派工单指的是每个成员在项目中的角色和职责，并未与每个工作包关联起来。责任分配矩阵不是6.5的输入； 选项B错误，项目章程不是6.5的输入； 选项C正确； 选项D错误，进度基准是6.5的输出，而非输入。



17、 [单选] 在规划某项目的进度计划时，项目经理注意到，项目章程中定义了一个不切实际的时间范围，而且某些里程碑预期会发生迟延，项目经理接下来该做什么？

When planning the schedule for a project, the project manager notices that the project charter defines an unrealistic timeframe and that certain milestones are expected to be delayed. What should the project manager do next?

A：遵守项目章程中规定的时间范围，因为这是一个业务需求

Comply with the time range specified in the project charter, as this is a business requirement

B：与有关的相关方开会解决这一偏差

Have a meeting with stakeholders to resolve this variance

C：调整时间范围，并寻求项目发起人批准

Adjust the time range and seek approval from the project sponsor

D：将该风险记录在风险日志中，并定期审查

Document this risk in the risk log and reviewed periodically

正确答案：B

你的答案：C

解析：解析：PMBOK（6）P207-6.5 需要在整个项目期间不断修订和维护项目进度模型，确保进度计划一直切实可行。 题干关键词“规划项目进度时”。 在进度基准被批准前，可以进行进度计划的更新，并无需遵循正式的变更流程。在这个时候，可以直接设法解决这一偏差，故选B； 选项A错误，项目章程不是6.5的输入； 选项C错误，项目经理无权修改项目章程中的时间范围； 选项D错误，这不是一个风险，而是一个确定会发生里程碑延期的问题。



21、 [单选] 一位经验丰富的项目经理加入团队，对一个陷入困境的项目执行健康检查。健康检查报告确定所有任务是同时执行的，并且从项目开始时就不断出现问题。若要提供更好的项目结果，原先的项目经理应该事先做什么？

An experienced project manager joins the team to execute a health check on a troubled project. The health check report confirmed that all tasks were performed at the same time, and issues continued to occur from the beginning of the project. To provide better project results, what should the former project manager have done beforehand?

A：审查详细的项目和阶段可交付成果

Review detailed project and phase deliverables

B：定义项目的制约因素和限制条件

Define project constraints and limitations

C：记录已识别的风险和假设条件

Document identified risks and assumptions

D：制定更详细的项目进度计划

Develop a more detailed project schedule

正确答案：D

你的答案：C

解析：解析：PMBOK（6）P205-6.5制定进度计划。 题干关键词“所有任务同时执行”。 由于所有任务同时执行并不断出现问题，可见原先的项目进度计划并没有创建合适的进度模型，故选D。 选项ABC在D之前，问“事先做什么”，选离问题时间点最近的那个选项。


